### 2013-06-30 Rcode/intro.rmd
corrected error introduced in diriKL
added text to describe the issue with the factor of 2 in local false sign rates; corrected bug in computation of localfdr with point mass

beta.ash=ash(ss$betahat,ss$betasd)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#compute the usual zscore and corresponding p value
zscore = ss$betahat/ss$betasd

@@

Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
```{r}
beta.ash$fitted.f
```

@@

>>>>

beta.ash=ash(ss$betahat,ss$betasd)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#these to test the VB version 
#beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
#beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)

#compute the usual zscore and corresponding p value
zscore = ss$betahat/ss$betasd

@@

Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
```{r}
beta.ash$fitted.f
beta.ash.auto$fitted.f
#beta.ash.vb.uniform$fitted.f
#beta.ash.vb.null$fitted.f
```

@@

### Some additional notes

#### Do we need a point mass at zero?

In some settings it is the convention to focus on testing whether $\beta_j=0$. However some dislike this focus, objecting that it is unlikely to be the case that $\beta_j=0$ exactly. For example, when comparing the average expression of a gene in human samples vs chimp samples, it might be considered unlikely that the expression
is *exactly* the same in both. Whether or not $\beta_j=0$
is considered unlikely may depend on the context.
However, in most contexts, finite data cannot
distinguish between $\beta_j=0$ and $\beta_j$ being very close to zero. Thus finite data cannot usually convince a skeptic that $\beta_j$ is exactly zero, rather than just very small. In contrast it is easy to imagine data that would convince a doubter that $\beta_j$ is truly non-zero. In this sense there is an assymetry between the inferences "$\beta_j$ is zero" and "$\beta_j$ is non-zero", an assymetry that is reflected in the admonition "failure to reject the null does not imply it to be true".

Thus any analysis that purports to distinguish between these cases must be making an assumption.

Consider two analyses of the same data, using two different "priors" $g$ for $\beta_j$, that effectively differ only in their assumptions about whether or not $\beta_j$ can be exactly zero. For concreteness, consider
\[ g_1(\cdot) = \pi \delta_0(\cdot) + (1-\pi) N(\cdot; 0,\sigma^2) \]
and
\[g_2(\cdot) = \pi N(\cdot; 0, \epsilon^2) + (1-\pi) N(\cdot; 0, \sigma^2).\]
If $\epsilon^2$ is sufficiently small, then these 
priors are "approximately the same", and will lead to "approximately the same" posteriors and inferences in many senses. To discuss these, let $p_j$ denote the posterior under prior $g_j$. Then, for any given (small) $\delta$, we will have $p_1(|\beta_j|<\delta) \approx p_2(|\beta_j|< \delta)$. However, we will not have $p_1(\beta_j=0) \approx p_2(\beta_j=0)$: the latter will always be zero, while the former could be appreciable.

 What if, instead, we examine $p_1(\beta_j >0)$ and $p_2(\beta_j >0)$? Again, these will differ. If this probability is big in the first analysis, say $1-\alpha$ with $\alpha$ small, then it could be as big as $1-\alpha/2$ in the second analysis. This is because if $p_1(\beta_j>0)=1-\alpha$, then $p_1(\beta_j=0)$ will often be close to $\alpha$, so for small $\epsilon$ $p_2(\beta_j)$ will have mass $\alpha$ near 0, of which half will be positive and half will be negative. 
Thus if we do an analysis without a point mass, but allow
for mass near 0, then we may predict what the results would have been if we had used a point mass.

Let's try: 
```{r}
beta.ash.pm = ash(ss$betahat, ss$betasd, auto=TRUE, usePointMass=TRUE)
beta.ash.pm$fitted.f
beta.ash.auto$fitted.f
plot(beta.ash.auto$localfdr,beta.ash.pm$localfdr,main="comparison of ash localfdr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=2)
```

Our conclusion: if we simulate data with a point mass,
and we analyze it without a point mass, we may underestimate
the false sign rate by a factor of 2. Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.
